---
title: "Final project"
author: "Yuxin Liu"
output: html_notebook
---
### Introduction
In this project, I'm going to compare and contrast 10 countries with highest happiness score and 10 counties with lowest score using the data achieved in 2015 and 2019 (because 2015 was the first year the survey system became relatively completed and the most recent data available was collected in 2019. Considering there is a 5-year gap between these two datasets, the result from the analysis would be convincing and we would be able to use some patterns we observed throughout the project to predict future trend. The datasets are available on World Happiness Report official website. The data was collected from the Gallup World Poll, and supported by the Ernesto Illy Foundation, illycaffè, Davines Group, Blue Chip Foundation, the William, Jeff, and Jennifer Gross Family Foundation, and Unilever’s largest ice cream brand Wall’s. The survey was conducted yearly after 2012, and the intention was to measure the happiness of people and to use the data to help guide public.  policy.

Guiding question: How did the factors impact on the top and bottom 10 happiest countries' happiness score in 2015 and 2019?




### Attributes intepretation     
Happiness_Rank: Rank of the country based on the Happiness Score        
Country: Name of the country    
Happiness_Score: A metric measured in 2015 and 2019 by asking the sampled people the question: "How would you rate your happiness"            
Economy_GDP_per_Capita: The extent to which GDP contributes to the calculation of the Happiness Score    
Social_Support: The extent to which Family contributes to the calculation of the Happiness Score           
Health_Life_Expectancy: The extent to which Life expectancy contributed to the calculation of the Happiness Score             
Freedom: The extent to which Freedom contributed to the calculation of the Happiness Score.         
Generosity: The extent to which Generosity contributed to the calculation of the Happiness Score.         
Trust_Government_Corruption: The extent to which Perception of Corruption contributes to Happiness Score.

### Data Access
```{r}
# clean up the RStudio environment 
rm(list = ls())

# Loading packages
library(DataComputing)
library(mosaic)
library(tidyverse)
library(lubridate)
library(ggplot2)
library(ggalt)
library(corrplot)
library(plotly)
library(gridExtra)
library(caTools)
library(rpart)
library(rpart.plot)
```



```{r}
# Data sources
df2015 <- read.csv("2015.csv")
df2019 <- read.csv("2019.csv")

# Data inspection
head(df2015, 10)  
head(df2019, 10)

summary(df2015)
summary(df2019)

nrow(df2015)
ncol(df2015)
nrow(df2019)
ncol(df2019)
```
Observation: There are 158 entries and 12 columns in df2015; 156 entries and 9 columns in df2019.     
Used 'summary' function to get general statistics, such as mean, median, minimum and maximum score. 

I noticed that although the number of rows and columns are not the same in df2015 and df2019, there are not many differences, and I'm going to select the features they both have for future analysis.     
Features appear in both tables: happiness rank, country, happiness score, GDP, family/social support, Healthy life expectancy, freedom, generosity, and perception of corruption.


### Data Wrangling
```{r message=FALSE, warning=FALSE}
# Create column 'Year'
Year <- c(2015)
Year2015 <- data.frame(Year, df2015)
# Rename the columns to formalize 
names(Year2015) <- c("Year", "Country", " Region", "Happiness_Rank","Happiness_Score","Standard_Error",  "Economy_GDP_per_Capita","Social_Support","Health_Life_Expectancy","Freedom","Trust_Government_Corruption", "Generosity", "Dystopia_Residual")


Year2 <- c(2019)
Year2019 <- data.frame(Year2, df2019)
names(Year2019) <- c("Year","Happiness_Rank","Country","Happiness_Score","Economy_GDP_per_Capita","Social_Support","Health_Life_Expectancy","Freedom","Generosity","Trust_Government_Corruption")

New2015 <- Year2015 %>%
  select("Happiness_Rank","Country","Happiness_Score","Economy_GDP_per_Capita","Social_Support","Health_Life_Expectancy","Freedom","Generosity","Trust_Government_Corruption", "Year")

# convert character to numeric
# combine data from 2015 and 2019 for future use
FinalTable <- rbind(New2015, Year2019, by= "Country")
FinalTable[3:10] <- lapply(FinalTable[3:10], as.numeric)
head(New2015)
head(Year2019)
FinalTable <- drop_na(FinalTable)
data2015 <- FinalTable %>%
  filter(Year=="2015")

data2019 <- FinalTable %>%
  filter(Year=="2019")

```



### Data Visualization
#### Overall correlation among variables in 2015 and 2019
```{r}
cormatrix2015 <- cor(data2015[3:9],use="pairwise.complete.obs")
corrplot(cormatrix2015, method = "number",tl.cex = 0.75, title="Correlation between Happiness scores and other attributs 2015",mar=c(0,0,1,0))

cormatrix2019 <- cor(data2019[3:9],use="pairwise.complete.obs")
corrplot(cormatrix2019, method = "number",tl.cex = 0.75, title="Correlation between Happiness scores and other attributs 2019",mar=c(0,0,1,0))

```

```{r}
Overall <- data.frame(x0 = c(2015,2015,2015,2015,2015,2015),    
                       y0 = c(0.78,0.74,0.72,0.57,0.18,0.4), 
                       x1 = c(2019,2019,2019,2019,2019,2019), 
                       y1 = c(0.79,0.78,0.78,0.57, 0.08, 0.39)) 
Overall$Attributes <- c('Economy_GDP_per_Capita','Social_Support','Health_Life_Expectancy','Freedom', 'Generosity', 'Trust_Government_Corruption ')

ggplot(Overall) + 
    geom_segment(aes(x = x0,y = y0,xend = x1,yend = y1,colour = Attributes)) + ggtitle("Overall relationship between attributes and happiness score \n changed from 2015 to 2019") + xlab("Year") + ylab("Correlation")
```
According to the graph, we can see that, as time goes by, the correlation between generosity and happiness score decreases, meaning that generosity doesn't impact on people's happiness as much as before. Meanwhile, the positive relationship between health life expectancy and happiness score, and social support and happiness score are increasing.

Side-by-side boxplot to show how factors related to happiness score have changed over time:

```{r}
par(mfrow=c(2,3))      
boxplot(Economy_GDP_per_Capita~Year,data=FinalTable)
boxplot(Social_Support~Year,data=FinalTable)
boxplot(Health_Life_Expectancy~Year,data=FinalTable)
boxplot(Freedom~Year,data=FinalTable)
boxplot(Generosity~Year,data=FinalTable)
boxplot(Trust_Government_Corruption~Year,data=FinalTable)
```

```{r}
# plot the density of happiness score for 2015 and 2019
ggplot(data2015, aes(x = Happiness_Score, fill="2015")) + geom_density()+ geom_density(data=data2019, aes(x=Happiness_Score, fill="2019"))
```



Now, let's list 10 happiest and 10 unhappiest countries in 2015 and 2019.

```{r}
#top 10
Top2015 <- data2015 %>% head(n=10)
Top2019 <- data2019 %>% head(n=10)


Top2015Graph <- ggplot(Top2015,aes(x=factor(Country, levels=Country), y=Happiness_Score)) + geom_bar(stat="identity", width = 0.5, fill="darkblue") +labs(title="10 Happiest Countries 2015", x="Country", y="Score")
Top2019Graph <- ggplot(Top2019,aes(x=factor(Country, levels=Country), y=Happiness_Score)) +geom_bar(stat="identity", width = 0.5, fill="darkgreen")+labs(title="10 Happiest Countries 2019", x="Country", y="Score")

# Bottom 10
Bottom2015 <- data2015 %>% tail(n=10)
Bottom2019 <- data2019 %>% tail(n=10)

Bottom2015Graph <- ggplot(Bottom2015,aes(x=factor(Country, levels=Country), y=Happiness_Score)) + geom_bar(stat="identity", width = 0.5, fill="darkblue") +labs(title="10 Least Happy Countries in 2015", x="Country", y="Score")
Bottom2019Graph <- ggplot(Bottom2019,aes(x=factor(Country, levels=Country), y=Happiness_Score)) +geom_bar(stat="identity", width = 0.5, fill="darkgreen")+labs(title="10 Least Happy Countries in 2019", x="Country", y="Score")

grid.arrange(Top2015Graph, Top2019Graph)
grid.arrange(Bottom2015Graph, Bottom2019Graph)
```

```{r}
# Mean of happiness score for top and bottom 10 happiest countries from 2015 to 2019
mean(Top2015$Happiness_Score)
mean(Bottom2015$Happiness_Score)

mean(Top2019$Happiness_Score)
mean(Bottom2019$Happiness_Score)


Change <- data.frame(x0 = c(2015,2015),    
                       y0 = c(7.4342, 3.3695), 
                       x1 = c(2019,2019), 
                       y1 = c(7.4559, 3.3041)) 
Change$Attributes <- c('mean of top 10','mean of bottom10')

ggplot(Change) + 
    geom_segment(aes(x = x0,y = y0,xend = x1,yend = y1,colour = Attributes)) + ggtitle("Mean of happiness score for top and bottom \n 10 happiest countries from 2015 to 2019") + xlab("Year") + ylab("Happiness Score")
```
The mean happiness score for the top 10 happiest countries has slightly improved as time goes by, while the mean of the bottom 10 happiest countries has slightly decreased. 

```{r}
cormatrix2015T <- cor(Top2015[3:9],use="pairwise.complete.obs")
corrplot(cormatrix2015T, method = "number",tl.cex = 0.6, title="Correlation between Happiness scores and \n other attributes for 10 happiest countries 2015",mar=c(0,0,2.5,0))

cormatrix2019T <- cor(Top2019[3:9],use="pairwise.complete.obs")
corrplot(cormatrix2019T, method = "number",tl.cex = 0.6, title="Correlation between Happiness scores and \n other attributes for 10 happiest countries 2019",mar=c(0,0,2.5,0))

cormatrix2015B <- cor(Bottom2015[3:9],use="pairwise.complete.obs")
corrplot(cormatrix2015B, method = "number",tl.cex = 0.6, title="Correlation between Happiness scores and \n other attributes for 10 least happy countries 2015",mar=c(0,0,2.5,0))

cormatrix2019B <- cor(Bottom2019[3:9],use="pairwise.complete.obs")
corrplot(cormatrix2019B, method = "number",tl.cex = 0.6, title="Correlation between Happiness scores and \n other attributes for 10 least happy countries 2019",mar=c(0,0,2.5,0))

```





```{r}
Happiest1 <- data.frame(x0 = c(2015,2015,2015,2015,2015,2015),    
                       y0 = c(0.53, 0.73,0.11,0.26, -0.43, -0.22), 
                       x1 = c(2019,2019,2019,2019,2019,2019), 
                       y1 = c(0.19, 0.69, -0.5, 0.57, -0.49, 0.28)) 
Happiest1$Attributes <- c('Economy_GDP_per_Capita','Social_Support','Health_Life_Expectancy','Freedom', 'Generosity', 'Trust_Government_Corruption ')

ggplot(Happiest1) + 
    geom_segment(aes(x = x0,y = y0,xend = x1,yend = y1,colour = Attributes)) + ggtitle("Relationship between attributes and happiness score for \n top 10 happiest countries changed from 2015 to 2019") + xlab("Year") + ylab("Correlation")

```

For top 10 happiest countries from 2015 to 2019:     

There's a positive relationship between social support and happiness score, but it has slightly decreased overtime.     
There's a positive relationship between GDP and happiness score but it has significantly dropped overtime.    
There's a positive relationship between freedom and happiness score, but it has increased overtime.     
There was a weak positive relationship between health life expectancy and happiness score when in 2015, but the correlation has become negative since 2016. On the contrast, the correlation between trust government corruption and happiness score has become positive in 2019 while it was negative in 2015.       
There's negative relationship between generosity and happiness score and it has slightly decreased as time goes by. 


```{r}
Least1 <- data.frame(x0 = c(2015,2015,2015,2015,2015,2015),    
                       y0 = c(0.12,0.63,-0.42,0.4,0.06,0.06), 
                       x1 = c(2019,2019,2019,2019,2019,2019), 
                       y1 = c(0.41,0.38,0.63,0.19,0.18,0.18)) 
Least1$Attributes <- c('Economy_GDP_per_Capita','Social_Support','Health_Life_Expectancy','Freedom', 'Generosity', 'Trust_Government_Corruption ')

ggplot(Least1) + 
    geom_segment(aes(x = x0,y = y0,xend = x1,yend = y1,colour = Attributes)) + ggtitle("Relationship between attributes and happiness score for \n bottom 10 happiest countries changed from 2015 to 2019") + xlab("Year") + ylab("Correlation")

```



```{r}
H_and_L_2015 <- data.frame(x0 = c("top10","top10","top10","top10","top10","top10"),    
                       y0 = c(0.53, 0.73,0.11,0.26, -0.43, -0.22), 
                       x1 = c("Bottom10","Bottom10","Bottom10","Bottom10","Bottom10","Bottom10"), 
                       y1 = c(0.12,0.63,-0.42,0.4,0.06,0.06)) 
H_and_L_2015$Attributes <- c('Economy_GDP_per_Capita','Social_Support','Health_Life_Expectancy','Freedom', 'Generosity', 'Trust_Government_Corruption ')

ggplot(H_and_L_2015) + 
    geom_segment(aes(x = x0,y = y0,xend = x1,yend = y1,colour = Attributes)) + ggtitle("Relationship between attributes and happiness score among \n top 10 and bottom 10 happiest countries in 2015") + xlab("Countries in 2015") + ylab("Correlation")
```





```{r}
H_and_L_2019 <- data.frame(x0 = c("top10","top10","top10","top10","top10","top10"),    
                       y0 = c(0.19, 0.69, -0.5, 0.57, -0.49, 0.28), 
                       x1 = c("Bottom10","Bottom10","Bottom10","Bottom10","Bottom10","Bottom10"), 
                       y1 = c(0.41,0.38,0.63,0.19,0.18,0.18)) 
H_and_L_2019$Attributes <- c('Economy_GDP_per_Capita','Social_Support','Health_Life_Expectancy','Freedom', 'Generosity', 'Trust_Government_Corruption ')

ggplot(H_and_L_2019) + 
    geom_segment(aes(x = x0,y = y0,xend = x1,yend = y1,colour = Attributes)) + ggtitle("Relationship between attributes and happiness score among \n top 10 and bottom 10 happiest countries in 2019") + xlab("Countries in 2019") + ylab("Correlation")
```
In 2019, there is a positive relationship between Health_Life_Expectancy and Happiness score for the bottom 10 countries, whereas the relationship between Health_Life_Expectancy and Happiness score for the top 10 countries is negative. To specify, in 2019, for bottom 10 countries, the higher the Health_Life_Expectancy is, the higher the happiness score will be. For top 10 countries, the higher health_Life_Expectancy matches with lower happiness score. Similarly, the correlation between generosity and happiness score for bottom 10 happiest countries is positive while that of top 10 countries in 2019 is negative.



### Machine learning models

```{r}
#install.packages('caTools')
set.seed(200)
dataset <- FinalTable[3:10]
split = sample.split(dataset$Happiness_Score, SplitRatio = 0.8)
training_set = subset(dataset, split == T)
test_set = subset(dataset, split == F)
dataFit = lm(formula = Happiness_Score ~ ., data = training_set)

summary(dataFit)
a <- predict(dataFit,dataset,interval="confidence",level=0.95)
plot(a)
anova(dataFit)

```

The equation of the multiple linear regression line
$\hat{HappinessScore}$=$122.43808 + 0.78789EconomyGDPperCapita + 1.17843SocialSupport +0.92656HealthLifeExpectancy + 1.66207Freedom + 0.24526Generosity + 0.8512TrustGovernmentCorruption - 0.05973Year$         

$R^2=0.7749$, meaning about 77.49% of the variation in happiness score is captured by the regression line based on the 6 variables we chose above.








```{r}
PredYFit = predict(dataFit, newdata = test_set)

PredActualFit <- as.data.frame(cbind(Prediction = PredYFit, Actual = test_set$Happiness_Score))

FitGraph <- ggplot(PredActualFit, aes(Actual, Prediction)) +
  geom_point() + theme_bw() + geom_abline() +
  labs(title = "Multiple Linear Regression", x = "Actual happiness score", y = "Predicted happiness score")
FitGraph
```




```{r}
# Fitting Decision Tree Regression
TreeFit = rpart(formula = Happiness_Score ~ .,data = dataset,control = rpart.control(minsplit = 20))

PredTreeFit = predict(TreeFit, newdata = test_set)

PredActualTree <- as.data.frame(cbind(Prediction = PredTreeFit, Actual = test_set$Happiness_Score))


TreeGraph <- ggplot(PredActualTree, aes(Actual, Prediction )) +
  geom_point() + theme_bw() + geom_abline() +
  labs(title = "Decision Tree Regression", x = "Actual happiness score", y = "Predicted happiness score")
TreeGraph
prp(TreeFit)
```

From the decision tree regression, we noticed that the decision tree won't be a good fit. Choose multiple linear regression model instead. 



### Conclusion
From the overall relationship between attributes and happiness score from 2015 to 2019, the correlation between happiness score and generosity and trust government corruption is decreasing overtime, while the correlation between happiness score and social support and health expectation has increased. Overall, GDP, social support, and health expectation are the top three attributes impact on happiness score the most.     



For top 10 happiest countries from 2015 to 2019, social support has always been the one that impact on happiness score the most. The positive relationship between freedom and happiness score has become stronger, same as trust government corruption and happiness score. GDP has become much less important for people's happiness.      
For bottom 10 happiest countries from 2015 to 2019, health life expectancy has become the most important attribute affects people's happiness. Relatively, freedom and social support became less important, which are opposite to that of top 10 happiest countries.        

In 2015, there's a stronger positive relationship between GDP and happiness score for top 10 happiest countries than that of bottom 10 happiest countries. On the opposite, the positive relationship between freedom and happiness score for top 10 happiest countries is weaker than that of bottom 10 happiest countries. When the correlation between health life expectancy and happiness score for top 10 happiest countries is positive, that of bottom 10 happiest countries are negative. Social support has always been the attribute that has the strongest positive relationship with happiness score for both top and bottom 10 happiest countries.     


In 2019, there's strong negative correlation between generosity and happiness score and health life expectancy and happiness score for top 10 happiest countries, while the correlation for bottom 10 happiest countries are positive and quite strong. The positive relationship between GDP and happiness score for bottom 10 countries are stronger than that of top 10 countries. 